---
title: "Demo - Parameterized R markdown"
author: "Nitesh Turaga"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        number_sections: true
params:
    geneA: "BRCA"
    geneB: "TP53"
    cancer_types: [Bladder Breast Liver]
    cancer_types_vector: !r as.character(c("Bladder", "Prostate", "Lung", "Breast"))
    my_vec: [1,2,3,4]
    int_value: !r as.integer(1000)
    logical_value: TRUE
    start: !r as.Date("2015-01-01")
---

This chuck below means `echo = TRUE` for every chunk

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# Parameters for this markdown file

```{bash, eval=FALSE}
---
params:
    ## character vector (length 1)
    geneA: "BRCA" 
    ## character vector (length 1)
    geneB: "TP53" 
    ## character vector (length 3)
    cancer_types: [Bladder Breast Liver] 
    ## character vector as an R object with !r
    cancer_types_vector: !r as.character(c("Bladder", "Prostate", "Lung", "Breast")) 
    ## numeric vector (length 4)
    my_vec: [1,2,3,4] 
    ## integer 
    int_value: !r as.integer(1000)
    ## logical value
    logical_value: TRUE
    ## Date object 
    start: !r as.Date("2015-01-01")
---
```


# Parameter types

Standard R types that can be parsed by the `yaml::yaml.load()`
function are supported

- `character`

- `integer`

- `numeric`

- `logical`

In addition, you can use arbitrary other R object types by specifying
the value using an R expression using `!r`.

Refer to this document - [Parameters][].

# Using Parameters in your text and code

## Character

You can refer to parameters in your chunks with `params$geneA`, call the
parameter with variable name that makes most sense for the parameter(not just `geneA`).

Within the text if you want to call on the first gene: `r params$geneA` and the second gene: `r params$geneB`, you can refer to them within your text with the back ticks and `r` call. This way you can make your paragraphs more modular and changeable.

```{bash eval=FALSE}
`r params$geneB`
```

Refer to [inline code][]

Reference your character parameters easily and reuse the variables

```{r}
(first_gene <- params$geneA)

(second_gene <- params$geneB)
```

## Vectors - character vector, integer vector etc

Use your vectors within code as well as needed: You can declare vectors 
in different ways

```
[Cancer1 Cancer2 Cancer3]
```

or 

```{bash eval=FALSE}
!r as.character(c('Cancer1', 'Cancer2', 'Cancer3'))
```

Example:

```{r}
## First type of cancer vector (length 1)
## cancer_types: [Bladder Breast Prostate] 
(cancer_types_vector1 <- params$cancer_types)
class(cancer_types_vector1)
unlist(strsplit(cancer_types_vector1, split = " "))

## second type of cancer vector 
## cancer_types_vector: !r as.character(c("Bladder", "Prostate", "Lung", "Breast"))
(cancer_types_vector2 <- params$cancer_types_vector)
class(cancer_types_vector2)
(params$cancer_types_vector[2])

## Integer vector (length 1)
(my_int_value <- params$int_value)

## Integer vector
(my_int_vector <- params$my_vec)
class(my_int_vector)
```


## Logical values

```{r}
## params:
##         logical_value: TRUE
(my_logical_value <- params$logical_value)
```

# Logical expressions using a logical value

Loop over your parameters to create great plots, IF your logical value is true

```{r}
## If param is TRUE
if (params$logical_value) {
    message("You will see something amazing!")
    for (cancer in params$cancer_types_vector) {
        print(cancer)
    }
}
```

Return some common cancers in both vectors, but don't print it out

```{r echo=TRUE}
## Return common cancers between two vectors
index <- params$cancer_types_vector %in% 
    unlist(strsplit(params$cancer_types, split = " ")) 
    

common_cancers <- params$cancer_types_vector[index]
```

These are the common cancer types - **`r common_cancers`** in both the vectors that were passed into the document as a parameters.

# Render outside of your rmarkdown

From outside of your Rmarkdown document you are able to render with
different parameters than the ones set inside your document, i.e,
re-use old Rmarkdowns and generate for different clients.

```{r eval=FALSE}
renderMyDocument <- function(template = "demo.Rmd", genes, cancers) {
  rmarkdown::render(template, params = list(
    geneA = genes[1],
    geneB = genes[2],
    cancers = cancers
  ))
}
```

# logical checking to run chunks

```{r do_it}
.is_valid <- function() {
    ## Make this as complex as your heart desires
    TRUE
}
## execute code
do_it = .is_valid()
```

Next you can pass this in chunk options i.e `eval=do_it`

```{r eval=do_it}
print("See it works!")
```


# Render more paramters

There are [advanced docs][] that allow you to parameterize even more.

```
---
title: "My Document"
output: html_document
params:
  minimum: 
    label: "Minimum:"
    value: 100
    input: slider
    min: 0
    max: 1000
  region:
    label: "Region:"
    value: east
    input: select
    choices: [east, west, north, south]
  data: 
    label: "Input dataset:"
    value: results.csv
    input: file
---
```

# Get a free shiny app for setting your params

This is a VERY VERY COOL FEATURE!!

```{r eval=FALSE}
rmarkdown::render("demo.Rmd", params = "ask")
```

# SessionInfo

```{r}
sessioninfo::session_info()
```


^[Inline code https://rmarkdown.rstudio.com/lesson-4.html]
^[advanced docs https://garrettgman.github.io/rmarkdown/developer_parameterized_reports.html]
^[Best book on Rmarkdown https://bookdown.org/yihui/rmarkdown/parameterized-reports.html]

[Parameters]: (https://rmarkdown.rstudio.com/lesson-6.html)
[inline code]: (https://rmarkdown.rstudio.com/lesson-4.html)
[advanced docs]: https://garrettgman.github.io/rmarkdown/developer_parameterized_reports.html       
